Refinements for Open Automata

### 2023-09-07 2021-Refinement/paper/refinementarticleV2.tex
More open transition consistency
vdash -> models

 We suppose given a satisfiability relation on closed formulas, denoted \({\vdash} f\).
% In practice one of our objective is to be able to use a SMT solver to reason automatically on the properties of open automata, in this case
%\(\vdash\) can hence be interpreted as an indicator of what is given to the SMT; it separates the external logic and the logic on \(\formulas\).
We will use two variants of the satisfiability relation:
\begin{itemize}
\item The satisfiability of a formula \(f \in \formulas\) under some valuation \(\sigma: V \to \values\) is defined as follows:
\( \sigma \vdash f \iff \vdash \exists \fvars{f\psubst{\sigma}}, f\psubst{\sigma} \)
\item The satisfiability of a formula \(f \in \formulas\) with some variable set \(V\) as context is defined as follows:
\( V \vdash f \iff  \vdash \forall V, \exists\mpar{\fvars{f} \setminus V}, f \)
\end{itemize}

@@

\subsection{Relations between Open Automata}
Establishing semantic equivalences and simulation relations between different OA requires  to compare their states. For this purpose, we suppose that the variables of the two OA are disjoint (a renaming of variables may have to be applied before comparing OA states).
\begin{definition}[Relation on open automata configurations] Suppose $V_1$ and $V_2$ are disjoint.
A relation on configurations of two OA \(\OAg[1]\) and \(\OAg[2]\) is a function \( \mathcal{R}: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\).
\end{definition}
%\TODO{C'est etrange, la formulation "the variables they refer to", comme il n'y a plus de variables d'etat... +1}
The idea is that two states are related depending on the satisfiability of the expression relying their variables, i.e., if the variables of the OA verify a certain formula. 
In other words, to each pair of states is attached a boolean formula that may refer to the  variables of each of the two OA, stating whether the two states are related or not.
Additionally, we say that the relation $ \mathcal{R}$ \emph{relates the initial states} of the automata if: \(\sigma_{01} \uplus \sigma_{02} \vdash  \mathcal{R}\mpar{s_{01}, s_{02}}\).
We illustrate such a relation over automata with bisimulation relation below.

%\TODO{I do not think we reason on configurations for the moment, if we want to, we should introduce this:}
%Two configurations \(\mpar{s_1, \sigma_1} \in S_1 \times\mpar{V_1 \to \values}, \mpar{s_2, \sigma_2} \in S_2 \times \mpar{V_2 \to \values}\) are related iff \(\sigma_1 \uplus \sigma_2 \vdash R\mpar{s_1, s_2}\).

@@

  \draw (1) edge[pos=.4] node[below=8pt]{\OTd{\alpha(x)}{h\mapsto\beta(x)}[x<0]{y\leftarrow -x}}  (2);
  \draw (1) edge[pos=.4] node[above]{\OTd{\alpha(x)}{h\mapsto\beta(x)}[x\geq 0]{y\leftarrow x}} (3);
\end{tikzpicture}
\hfill
\begin{tikzpicture}
  \node[state,inner sep=5pt,minimum size=10pt] (1) {$s_2$};
  \node[state,inner sep=5pt,minimum size=10pt,  right of=1,right=90pt] (2) {$t_2$};
\draw (1) edge[align=center] node[below]{\OTd{\alpha(x)}{h\mapsto\beta(x)}{z\leftarrow x}}  (2);
%  \draw (1) edge[align=center] node[above]{\OTd{\alpha(x)}{h\mapsto\beta(x)}{x\geq 0}{y\leftarrow x}} (3);
\end{tikzpicture}~

@@

%\begin{comment}
\subsection{Reachability}
We finally define a new predicate abstracting state reachability for OA, it  allows us to reason on reachable states in an automaton. It can  be seen as an abstraction of the reachable states under the form of a predicate that must stay verified along the execution of the OA.
\begin{definition}[Reachability] \label{Def:Reach}
For any OA \(A = \OAg\), a reachability predicate \(\reach{A}: S \to \rformulas[V]\) is any predicate on states that is valid on initial state, and preserved across transitions:
\[\sigma_0 \vdash \reach{A}\mpar{s_0}\quad\land\quad\forall t = \OTg \in T, \fvars{t} \vdash \left(\reach{A}\mpar{s} \wedge g \implies \reach{A}\mpar{s'}\psubst{\psi}\right)\]
\end{definition}

@@

%
%\begin{definition}[Deadlock reduction]\label{def:dpwd}\\
%Let \(\OAg[1]\) and \(\OAg[2]\) be two OAs.
%A relation on OA configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is deadlock reducing if  it satisfies the following\footnote{Note that variables of $T_1$ are existentially quantified in the proposition.}:
%\begin{multline*}
% \forall \mpar{s_1, s_2} \in S_1 \times S_2,
% \\ V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash \left(R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \right)
%\end{multline*}
%\end{definition}

@@

%
%\begin{definition}[Non-blocking composition]\label{Def:Non-blockcomp}
%Let \(A_i = \OAg[i]\) with \( 0 \leq i \leq n\) be a family of open automata.
%Let \(A = A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\) and $A= \OAg$.
%
%The composition \(A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\) is non-blocking if   \(A \) has a reachability predicate such that, for each reachable configuration, if there is a possible transition in \(A_0\) then there is a possible transition in \(A\):
%\[ \forall s = \mpar{s_0, s_1,..,s_n} \in S, V \uplus \biguplus_\subbox{t_0 \in \fOT{s_0}} \fvars{t_0} \vdash \left(\reach{A}\mpar{s} \wedge \bigvee_\subbox{t_0 \in \fOT{s_0}} \fguard{t_0} \implies \bigvee_\subbox{t \in \fOT{s}} \fguard{t} \right)\]
%\end{definition}
%Again we use guards to ensure that the transition can occur. It is not sufficient to ensure the existence of equivalent transitions in general, but it will be sufficient in the context of refinement.

@@

However here we want to build a simulation relation that  also guarantees that no deadlock is introduced when refining the automaton. This property is quite frequent in simulation relation, and referred to as \emph{lack of new deadlocks}~\cite{Kouchnarenko:2007} or \emph{complete simulation}~\cite{sangiorgi:bisim-coind12}.
The notion of deadlock should however be specialised to our OA. 
Indeed, it is not very useful to check the  existence of a transition, instead it makes more sense to use the guards to check if a transition can be taken.
 We thus define a deadlock reduction criterion based on how the outgoing transitions are guarded.
As such, a simulation does not introduce deadlocks if in the conditions where no transition is possible in the refined automaton, no transition were already possible in the more general one. More formally, for any pair of states $s_1$ and $s_2$  we introduce a criterion of the form: 
\begin{multline*}
\left(\forall \mpar{s_1, s_2} \!\in\! S_1 \!\times\! S_2, V_1 \uplus V_2  \vdash \mathcal{R}\mpar{s_1, s_2} \wedge \lnot( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1})\right) \\\implies \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  )
\end{multline*}

Which can be rewritten as:
\[
\left(
\forall \mpar{s_1, s_2} \!\in\! S_1 \!\times\! S_2, V_1 \uplus V_2  \vdash \mathcal{R}\mpar{s_1, s_2}\right) \implies( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} ) \lor¢ \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  )
\]

Both statements being equivalent, as each of them may reveal more intuitive than the other in different situations, we  use them interchangeably.
%Let us first explain the reasons why simulation can introduce allow undesired behaviours to emerge and how we want to limit them. 
%The automaton that simulates a specification features any behaviour that contains the behaviour of the specification, or more precisely, for each state that simulates a state of the specification, the simulating automaton must at least feature the behaviour of the specification.
% In this definition, nothing prevents the implementation from 
%
%adding non-determinism and non-deterministically reaching a state that is unrelated to the specification, this state could even be a deadlock even though we only followed transitions that exist in the specification.
%To avoid the appearance of such deadlocks, we will state that, when both the implementation and the specification do the same transition, they stay related.
%This ensures a form of continuity in the comparison in the sense that the refined system is constrained to follow the behaviour of the simulated system along a given scenario. 
%In terms of traces this amounts to some form of deadlock reduction: for each trace followed by the simulated system if the refined system followed the beginning of the trace, it must be able to keep following the trace. 
%Of course, the classical trace view is a simplified view compared to the guards, states and actions from  the holes that exist in OA, but it gives the right intuition.
%
%Next definition states that if $T_2$ is not in a deadlock position then $T_1$ can do a transition but not necessarily with the same input values. This is very weak but sufficient when composed with the def  of simulation.
%\begin{definition}[Preservation along identical transitions]\label{def:preservation} \\
%Let \(\OAg[1]\) and \(\OAg[2]\) be two OAs.
%A relation \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) between  \(\OAg[1]\) and \(\OAg[2]\)
%  is preserved along identical transitions if  it satisfies the following:
%\begin{multline*}
%\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} \mpar{t_1, t_2} = \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2} \\  V_1 \uplus V_2 \uplus \fvars{t_1} \uplus \fvars{t_2} \vdash \hspace{6cm}\\  R\mpar{s_1, s_2} \wedge 	g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \implies  R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}\\
%\end{multline*}
%\end{definition}
%We thus define the following property that can be understood as follows. Take a specification automaton and its implementation. Consider two related states for the refinement relation. Suppose an identical transition exist in both automata (with same guards and same label), then either the reached states $s_1'$ and $s_2'$ are related by refinement, or there exists another transition of the specification that is also identical and leads to a state $s_1''$ that is related to $s_2'$ in the refinement relation.
%Finally, we should take into account that 1) instead of equality, we should check if there is a state such that both transitions are feasible (the conjunction of guards is satisfiable), and that  2) instead of finding another transition of the specification we might find a family of transitions that cover all the cases of the transition of the implementation. 
We can now state the definition of simulation between OA that have the same set of holes.

\begin{definition}[Hole-equal simulation]\label{def:HoleEqualSim}
Consider two OA with identical set of holes:
\(A_1 = \OA{S_1}{s_{01}}{V_1}{\sigma_{01}}{J}{T_1}\) and \(A_2 = \OA{S_2}{s_{02}}{V_2}{\sigma_{02}}{J}{T_2}\), the relation on configurations \(\mathcal{R}: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a hole-\(\symb{equal}\) simulation from $A_1$ to $A_2$ if the following conditions hold : 
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \vdash \mathcal{R}\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)\vspace{-8pt}
\noindent\begin{multline*}
 \bigsymb{\forall} {t_1} = \openrule
         {
           \beta_{1 j}^{j\in J'_{1}}, g_1,\psi_1}
         {s_1 \OTarrow {\alpha_1} s_1'} \in \fOT{s_1} .\,~~
\bigsymb{\bigexists} 
\mpar{t_{2x} \!=\! \openrule
         {
           \beta_{2 x j}^{j\in J'_{2x}}, g_{2 x},\psi_{2 x}}
         {s_2 \OTarrow {\alpha_{2 x}} s_{2 x}'}
}^{x \in X}\in \fOT{s_2}.\,~~\\[2pt]
\mpar{\forall x \in X, J'_{2x} = J'_1 } \nwedge\hfill\\[1pt]
%\everymath{\displaystyle}\begin{array}{l}
V_1 \uplus V_2 \uplus  \fvars{t_1} \vdash 
 \mathcal{R}\mpar{s_1, s_2} \wedge g_1 \implies
 \operatorname*{\bigsymb{\bigvee}}_{x \in X}
{\everymath{\displaystyle}
\mpar{\begin{array}{l}
			\alpha_{2x} = \alpha_{1} \wedge \bigwedge_\subbox{j \in J'_1} \beta_{2xj} = \beta_{1j} \nwedge\\[12pt]
			 g_{2x} \wedge\mathcal{R}\mpar{ s'_{1},s'_{2x}}\psubst{\psi_{2x} \uplus \psi_{1}}
		\end{array}} }
%\end{array} 
\end{multline*}
\item [(3)] Deadlock reduction:
\[
\forall \mpar{s_1, s_2} \!\in\! S_1 \!\times\! S_2, V_1 \uplus V_2  \!\vdash\! \mathcal{R}\mpar{s_1, s_2} \implies( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} ) \lor¢ \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  )
\]
If there is a hole-equal simulation from $A_1$ to $A_2$, then we say that  $A_2$ simulates $A_1$; we  denote it $A_2\leq A_1$.
\end{definition}

@@

\begin{definition}[Hole-tracking simulation]\label{Def:OA-Refinement}
For two OA\\ \(A_1 \defobject \OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\) and \(A_2 \defobject \OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\), \(A_1\) is a simulation of \(A_2\) tracking holes \(H\), noted \(\wrel{A_1}{A_2}{H}\), with \(H \subseteq J_1 \cap J_2\), if there is a relation on configurations $\mathcal{R}: \mpar{S_1 \times S_2} \to \rformulas[V_1 \uplus V_2]$ such that\footnote{Note that the definition below is identical to the hole equal simulation except $\cap H$ is added in a few places.}:
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \vdash \mathcal{R}\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)
\begin{multline*}
	\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{\OTx{2}{x}{2x}{2x} \in \fOT{s_2}}^{x \in X}, \\[12pt]
		\quad \mpar{\forall x \in X, J'_{2x} \cap H = J'_1 \cap H} \nwedge\hfill\\[1pt]
		 V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\\hspace{5em} \mpar{\mathcal{R}\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_\subbox{j \in J'_{1} \cap H} \beta_{1j} = \beta_{2xj}   \nwedge\\
			 g_{2x} \wedge \mathcal{R}\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}}} 
	\end{array} 
\end{multline*}
\item[(3)] Deadlock reduction:
\[
\forall \mpar{s_1, s_2} \!\in\! S_1 \!\times\! S_2, V_1 \uplus V_2  \vdash \mathcal{R}\mpar{s_1, s_2} \implies( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} ) \lor¢ \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  )
\]
\end{definition}
Note that every action of the holes outside \(H\) is unconstrained according to the simulation relation.

@@

\TODO{remove this def}
\begin{definition}[Deadlock reduction]\label{def:dpwd}\\
Let \(\OAg[1]\) and \(\OAg[2]\) be two OAs.
A relation on OA configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is deadlock reducing if  it satisfies the following\footnote{Note that variables of $T_1$ are existentially quantified in the proposition.}:
\begin{multline*}
 \forall \mpar{s_1, s_2} \in S_1 \times S_2,\\ V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash \left(R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \right)
\end{multline*}
\end{definition}

>>>>

 We suppose given a satisfiability relation on closed formulas, denoted \({\models} f\).
% In practice one of our objective is to be able to use a SMT solver to reason automatically on the properties of open automata, in this case
%\(\models\) can hence be interpreted as an indicator of what is given to the SMT; it separates the external logic and the logic on \(\formulas\).
We will use two variants of the satisfiability relation:
\begin{itemize}
\item The satisfiability of a formula \(f \in \formulas\) under some valuation \(\sigma: V \to \values\) is defined as follows:
\( \sigma \models f \iff \models \exists \fvars{f\psubst{\sigma}}, f\psubst{\sigma} \)
\item The satisfiability of a formula \(f \in \formulas\) with some variable set \(V\) as context is defined as follows:
\( V \models f \iff  \models \forall V, \exists\mpar{\fvars{f} \setminus V}, f \)
\end{itemize}

@@

\subsection{Relations between Open Automata}
Establishing semantic equivalences and simulation relations between different OA requires  to compare their states. For this purpose, we suppose that the variables of the two OA are disjoint (a renaming of variables may have to be applied before comparing OA states).
\begin{definition}[Relation on open automata configurations] Suppose $V_1$ and $V_2$ are disjoint.
A relation on configurations of two OA \(\OAg[1]\) and \(\OAg[2]\) is a function \( \mathcal{R}: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\).
\end{definition}
%\TODO{C'est etrange, la formulation "the variables they refer to", comme il n'y a plus de variables d'etat... +1}
The idea is that two states are related depending on the satisfiability of the expression relying their variables, i.e., if the variables of the OA verify a certain formula. 
In other words, to each pair of states is attached a boolean formula that may refer to the  variables of each of the two OA, stating whether the two states are related or not.
Additionally, we say that the relation $ \mathcal{R}$ \emph{relates the initial states} of the automata if: \(\sigma_{01} \uplus \sigma_{02} \models  \mathcal{R}\mpar{s_{01}, s_{02}}\).
We illustrate such a relation over automata with bisimulation relation below.

%\TODO{I do not think we reason on configurations for the moment, if we want to, we should introduce this:}
%Two configurations \(\mpar{s_1, \sigma_1} \in S_1 \times\mpar{V_1 \to \values}, \mpar{s_2, \sigma_2} \in S_2 \times \mpar{V_2 \to \values}\) are related iff \(\sigma_1 \uplus \sigma_2 \models R\mpar{s_1, s_2}\).

@@

  \draw (1) edge[pos=.4] node[below=6pt]{\OTd{\alpha(x)}{h\mapsto\beta(x)}[x<0]{y\leftarrow -x}}  (2);
  \draw (1) edge[pos=.4] node[above]{\OTd{\alpha(x)}{h\mapsto\beta(x)}[x\geq 0]{y\leftarrow x}} (3);
\end{tikzpicture}
\hfill
\begin{tikzpicture}
  \node[state,inner sep=5pt,minimum size=10pt] (1) {$s_2$};
  \node[state,inner sep=5pt,minimum size=10pt,  right of=1,right=90pt] (2) {$t_2$};
\draw (1) edge[align=center] node[below]{\OTd{\alpha(x)}{h\mapsto\beta(x)}{z\leftarrow x}}  (2);
%  \draw (1) edge[align=center] node[above]{\OTd{\alpha(x)}{h\mapsto\beta(x)}{x\geq 0}{y\leftarrow x}} (3);
\end{tikzpicture}~

@@

%\begin{comment}
\subsection{Reachability}
We finally define a new predicate abstracting state reachability for OA, it  allows us to reason on reachable states in an automaton. It can  be seen as an abstraction of the reachable states under the form of a predicate that must stay verified along the execution of the OA.
\begin{definition}[Reachability] \label{Def:Reach}
For any OA \(A = \OAg\), a reachability predicate \(\reach{A}: S \to \rformulas[V]\) is any predicate on states that is valid on initial state, and preserved across transitions:
\[\sigma_0 \models \reach{A}\mpar{s_0}\quad\land\quad\forall t = \OTg \in T, \fvars{t} \models \left(\reach{A}\mpar{s} \wedge g \implies \reach{A}\mpar{s'}\psubst{\psi}\right)\]
\end{definition}

@@

%
%\begin{definition}[Deadlock reduction]\label{def:dpwd}\\
%Let \(\OAg[1]\) and \(\OAg[2]\) be two OAs.
%A relation on OA configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is deadlock reducing if  it satisfies the following\footnote{Note that variables of $T_1$ are existentially quantified in the proposition.}:
%\begin{multline*}
% \forall \mpar{s_1, s_2} \in S_1 \times S_2,
% \\ V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \models \left(R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \right)
%\end{multline*}
%\end{definition}

@@

%
%\begin{definition}[Non-blocking composition]\label{Def:Non-blockcomp}
%Let \(A_i = \OAg[i]\) with \( 0 \leq i \leq n\) be a family of open automata.
%Let \(A = A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\) and $A= \OAg$.
%
%The composition \(A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\) is non-blocking if   \(A \) has a reachability predicate such that, for each reachable configuration, if there is a possible transition in \(A_0\) then there is a possible transition in \(A\):
%\[ \forall s = \mpar{s_0, s_1,..,s_n} \in S, V \uplus \biguplus_\subbox{t_0 \in \fOT{s_0}} \fvars{t_0} \models \left(\reach{A}\mpar{s} \wedge \bigvee_\subbox{t_0 \in \fOT{s_0}} \fguard{t_0} \implies \bigvee_\subbox{t \in \fOT{s}} \fguard{t} \right)\]
%\end{definition}
%Again we use guards to ensure that the transition can occur. It is not sufficient to ensure the existence of equivalent transitions in general, but it will be sufficient in the context of refinement.

@@

However here we want to build a simulation relation that  also guarantees that no deadlock is introduced when refining the automaton. This property is quite frequent in simulation relation, and referred to as \emph{lack of new deadlocks}~\cite{Kouchnarenko:2007} or \emph{complete simulation}~\cite{sangiorgi:bisim-coind12}.
The notion of deadlock should however be specialised to our OA. 
Indeed, it is not very useful to check the  existence of a transition, instead it makes more sense to use the guards to check if a transition can be taken.
 We thus define a deadlock reduction criterion based on how the outgoing transitions are guarded.
As such, a simulation does not introduce deadlocks if in the conditions where no transition is possible in the refined automaton, no transition were already possible in the more general one. More formally, for any pair of states $s_1$ and $s_2$  we introduce a criterion of the form: 
\begin{multline*}
\left(\forall \mpar{s_1, s_2} \!\in\! S_1 \!\times\! S_2, V_1 \uplus V_2  \models \mathcal{R}\mpar{s_1, s_2} \wedge \lnot( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1})\right) \\\implies \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  )
\end{multline*}

Which can be rewritten as:
\[
\left(
\forall \mpar{s_1, s_2} \!\in\! S_1 \!\times\! S_2, V_1 \uplus V_2  \models \mathcal{R}\mpar{s_1, s_2}\right) \implies( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} ) \lor¢ \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  )
\]

Both statements being equivalent, as each of them may reveal more intuitive than the other in different situations, we  use them interchangeably.
%Let us first explain the reasons why simulation can introduce allow undesired behaviours to emerge and how we want to limit them. 
%The automaton that simulates a specification features any behaviour that contains the behaviour of the specification, or more precisely, for each state that simulates a state of the specification, the simulating automaton must at least feature the behaviour of the specification.
% In this definition, nothing prevents the implementation from 
%
%adding non-determinism and non-deterministically reaching a state that is unrelated to the specification, this state could even be a deadlock even though we only followed transitions that exist in the specification.
%To avoid the appearance of such deadlocks, we will state that, when both the implementation and the specification do the same transition, they stay related.
%This ensures a form of continuity in the comparison in the sense that the refined system is constrained to follow the behaviour of the simulated system along a given scenario. 
%In terms of traces this amounts to some form of deadlock reduction: for each trace followed by the simulated system if the refined system followed the beginning of the trace, it must be able to keep following the trace. 
%Of course, the classical trace view is a simplified view compared to the guards, states and actions from  the holes that exist in OA, but it gives the right intuition.
%
%Next definition states that if $T_2$ is not in a deadlock position then $T_1$ can do a transition but not necessarily with the same input values. This is very weak but sufficient when composed with the def  of simulation.
%\begin{definition}[Preservation along identical transitions]\label{def:preservation} \\
%Let \(\OAg[1]\) and \(\OAg[2]\) be two OAs.
%A relation \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) between  \(\OAg[1]\) and \(\OAg[2]\)
%  is preserved along identical transitions if  it satisfies the following:
%\begin{multline*}
%\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} \mpar{t_1, t_2} = \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2} \\  V_1 \uplus V_2 \uplus \fvars{t_1} \uplus \fvars{t_2} \models \hspace{6cm}\\  R\mpar{s_1, s_2} \wedge 	g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \implies  R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}\\
%\end{multline*}
%\end{definition}
%We thus define the following property that can be understood as follows. Take a specification automaton and its implementation. Consider two related states for the refinement relation. Suppose an identical transition exist in both automata (with same guards and same label), then either the reached states $s_1'$ and $s_2'$ are related by refinement, or there exists another transition of the specification that is also identical and leads to a state $s_1''$ that is related to $s_2'$ in the refinement relation.
%Finally, we should take into account that 1) instead of equality, we should check if there is a state such that both transitions are feasible (the conjunction of guards is satisfiable), and that  2) instead of finding another transition of the specification we might find a family of transitions that cover all the cases of the transition of the implementation. 
We can now state the definition of simulation between OA that have the same set of holes.

\begin{definition}[Hole-equal simulation]\label{def:HoleEqualSim}
Consider two OA with identical set of holes:
\(A_1 = \OA{S_1}{s_{01}}{V_1}{\sigma_{01}}{J}{T_1}\) and \(A_2 = \OA{S_2}{s_{02}}{V_2}{\sigma_{02}}{J}{T_2}\), the relation on configurations \(\mathcal{R}: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a hole-\(\symb{equal}\) simulation from $A_1$ to $A_2$ if the following conditions hold : 
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \models \mathcal{R}\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)\vspace{-8pt}
\noindent\begin{multline*}
 \bigsymb{\forall} {t_1} = \OTx{1}{}{1}{1} \in \fOT{s_1} .\,~~
\bigsymb{\bigexists} 
\mpar{t_{2x} \!=\! \OTx{2}{x}{2x}{2x}}^{x \in X}\in \fOT{s_2}.\,~~\\[2pt]
\mpar{\forall x \in X, J'_{2x} = J'_1 } \nwedge\hfill\\[1pt]
%\everymath{\displaystyle}\begin{array}{l}
V_1 \uplus V_2 \uplus  \fvars{t_1} \models 
 \mathcal{R}\mpar{s_1, s_2} \wedge g_1 \implies
 \operatorname*{\bigsymb{\bigvee}}_{x \in X}
{\everymath{\displaystyle}
\mpar{\begin{array}{l}
			\alpha_{2x} = \alpha_{1} \wedge \bigwedge_\subbox{j \in J'_1} \beta_{2xj} = \beta_{1j} \nwedge\\[12pt]
			 g_{2x} \wedge\mathcal{R}\mpar{ s'_{1},s'_{2x}}\psubst{\psi_{2x} \uplus \psi_{1}}
		\end{array}} }
%\end{array} 
\end{multline*}
\item [(3)] Deadlock reduction:
\[
\forall \mpar{s_1, s_2} \!\in\! S_1 \!\times\! S_2, V_1 \uplus V_2  \!\models\! \mathcal{R}\mpar{s_1, s_2} \implies( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} ) \lor¢ \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  )
\]
If there is a hole-equal simulation from $A_1$ to $A_2$, then we say that  $A_2$ simulates $A_1$; we  denote it $A_2\leq A_1$.
\end{definition}

@@

\begin{definition}[Hole-tracking simulation]\label{Def:OA-Refinement}
For two OA\\ \(A_1 \defobject \OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\) and \(A_2 \defobject \OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\), \(A_1\) is a simulation of \(A_2\) tracking holes \(H\), noted \(\wrel{A_1}{A_2}{H}\), with \(H \subseteq J_1 \cap J_2\), if there is a relation on configurations $\mathcal{R}: \mpar{S_1 \times S_2} \to \rformulas[V_1 \uplus V_2]$ such that\footnote{Note that the definition below is identical to the hole equal simulation except $\cap H$ is added in a few places.}:
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \models \mathcal{R}\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)
\begin{multline*}
	\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{\OTx{2}{x}{2x}{2x} \in \fOT{s_2}}^{x \in X}, \\[12pt]
		\quad \mpar{\forall x \in X, J'_{2x} \cap H = J'_1 \cap H} \nwedge\hfill\\[1pt]
		 V_1 \uplus V_2 \uplus \fvars{t_1} \models\\\hspace{5em} \mpar{\mathcal{R}\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_\subbox{j \in J'_{1} \cap H} \beta_{1j} = \beta_{2xj}   \nwedge\\
			 g_{2x} \wedge \mathcal{R}\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}}} 
	\end{array} 
\end{multline*}
\item[(3)] Deadlock reduction:
\[
\forall \mpar{s_1, s_2} \!\in\! S_1 \!\times\! S_2, V_1 \uplus V_2  \models \mathcal{R}\mpar{s_1, s_2} \implies( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} ) \lor¢ \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  )
\]
\end{definition}
Note that every action of the holes outside \(H\) is unconstrained according to the simulation relation.

@@

\TODO{remove this def}
\begin{definition}[Deadlock reduction]\label{def:dpwd}\\
Let \(\OAg[1]\) and \(\OAg[2]\) be two OAs.
A relation on OA configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is deadlock reducing if  it satisfies the following\footnote{Note that variables of $T_1$ are existentially quantified in the proposition.}:
\begin{multline*}
 \forall \mpar{s_1, s_2} \in S_1 \times S_2,\\ V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \models \left(R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \right)
\end{multline*}
\end{definition}
